---
title: "PH142 Midterm II Review"
output: html_notebook
---

<b>Note to fellow GSI's</b>: Some of this does not need to be read aloud, but rather just as some extra conceptual material for students to read on their own.

This review is not exhaustive. Please consult lecture notes and your text. Good luck studying to all!

## Experimental Design

### Chapter 6

### Chapter 7

## Probability

### Chapter 9

### Chapter 10

## Distributions
A common question has been <i><b>WHEN DO WE USE THESE DISTRIBUTIONS!?</i></b> Well, sometimes our data tell a story that we see over and over again. We can think of the typical Korean drama or the common pop song.  

<u>Recipe for the Typical Korean Drama</u>
<ol>
  <li>Story revolves around 2 potential lovers</li>
  <li>One of the lovers is super dopey/clumsy or nonconforming to some standard of beauty</li>
  <li>The other lover is mean</li>
  <li>Many dramatic, slow motion moments betweent the 2 lovers</li>
  <li>After many tribulations, the lovers end up as a couple</li>
</ol>

<u>Recipe for a LOT of Pop Songs</u>
<ol>
  <li>Chords: E B C#m A</li>
  <li>Verse 1, Verse 2, Chorus, Bridge, Chorus or some very similar format</li>
</ol>

We can think of distributions as things that model patterns in data. If we see data that looks approximately normal or follows a recipe for the Binomial or Poisson(*), then we should definitely consider using the appropriate distribution. When we use a distribution like one of the named three above, we can end up saying a lot about our data. We can say the mean, the variance/standard deviation, and calculate probabilities.

(*) See the following Chapter 11 and 12 notes for these recipes.

### Chapter 11: Normal Distribution
INSERT PICTURE OF NORMAL DISTRIBUTION

<u>Recipe for Normal</u>  
This isn't quite a recipe, but more like guidelines.
<ol>
  <li>Goals: We want to find a probability, i.e. $P(X>k), P(X<k), P(X \geq K), P(X \leq k)$ but <i>NEVER</i> $P(X=k)$</li>
  <li>The data are said to be normal or are assumed to be normal in the prompt</li>
</ol>

##### Question
What is the difference between normal and standard normal?

##### Example
According to <a href="https://www.basketball-reference.com/leagues/NBA_stats.html">basketball-reference.net</a>, the mean height for NBA players is 6'7" = 79 inches. Suppose it is known that player height is normally distributed with a standard deviation of 4 inches. What is the probability of a ball player to be shorter than 6 feet?

The prompt information translates to:
```{r}
mu  <- 79
sd  <- 4
k   <- 6*12
```

We can calculate this probability one way (without tables!).
```{r}
pnorm(q=k, mean=mu, sd=sd)
```

All these `[ ]norm` functions are fair game!

### Chapter 12: Binomial and Poisson Distributions

#### Binomial
INSERT BINOMIAL DISTRIBUTION HERE

<u>Recipe for Binomial</u>
<ol>
  <li>Goals: We want to find a probability, these are all fine: $P(X>k), P(X<k), P(X \geq K), P(X \leq k), P(X=k)$</li>
  <li>You have some fixed amount of trials</li>
  <li>Trials are independent</li>
  <li>The probability of success is the same for each trial</li>
  <li>Assumption k</li>
</ol>

##### Question
When can we approximate the binomial as a normal distribution? Why?

##### Example
A shopper goes online Thursday mornings to attempt to purchase Supreme apparel and accessories. Due to the brand's popularity and limited supply, the shopper successfully purchases any item 1 in 10 times. What is the probability that the shopper will make 4 successful purchases after 10 independent purchase attempts? How about greater than 4?

The prompt information translates to:
```{r}
n_trials    <- 10
k_success   <- 4
probability <- 1/10
```

And we can calculate the first quantity "by hand" using the formula:
```{r}
choose(n_trials, k_success)*(probability)^k_success*(1-probability)^(n_trials-k_success)
```

Or we can use a special function:
```{r}
dbinom(x=k_success, size=n_trials, prob=probability)
```

The second quantity can be found by using the Complement Rule.
```{r}
1 - choose(n_trials, 4)*(probability)^(4)*(1-probability)^(n_trials-4) - choose(n_trials, 3)*(probability)^(3)*(1-probability)^(n_trials-3) - choose(n_trials, 2)*(probability)^(2)*(1-probability)^(n_trials-2) - choose(n_trials, 1)*(probability)^(1)*(1-probability)^(n_trials-1) - choose(n_trials, 0)*(probability)^(0)*(1-probability)^(n_trials-0)
```

Or using another cool function that calculate the sum of probabilities from 0 to k.

```{r}
1-pbinom(q=k_success, size=n_trials, prob=probability)
```

Or even more conveniently:
```{r}
pbinom(q=k_success, size=n_trials, prob=probability, lower.tail=FALSE)
```

All these `[ ]binom` functions are fair game!

#### Poisson
INSERT POISSON DISTRIBUTION HERE

<u>Recipe for Poisson</u>
<ol>
  <li>Goals: We want to find a probability, these are all fine: $P(X>k), P(X<k), P(X \geq K), P(X \leq k), P(X=k)$</li>
  <li>The data are said to be normal or are assumed to be normal in the prompt</li>
  <li>Assumption 1</li>
</ol>

##### Example
San Francisco is known for its fog. In fact, the fog's name is Karl! We expect a warm, sunny, Karl-free day in San Francisco about twice per month.

All these `[ ]pois` functions are fair game!

## Sampling Distributions

### Chapter 13